Accessible, realistic genome simulation with selection using \stdpopsim

### 2025-03-19 analysis2_manuscript.tex
Ilan's final round of edits (#60)
* Ilan's edits:
- marked with %IG non-trivial changes
- removed previous resolved comments

* one more tweak

* Update analysis2_manuscript.tex

Co-authored-by: Andrew Kern <adkern@uoregon.edu>

* Update analysis2_manuscript.tex

Co-authored-by: Andrew Kern <adkern@uoregon.edu>

* Update analysis2_manuscript.tex

Co-authored-by: Andrew Kern <adkern@uoregon.edu>

* Update analysis2_manuscript.tex

Co-authored-by: Andrew Kern <adkern@uoregon.edu>

* Update analysis2_manuscript.tex

* Update analysis2_manuscript.tex

---------

Co-authored-by: Ilan Gronau <gronau

% author tranche explanation:
% co-first authors
\author[1,*]{Graham Gower}
\author[2,*]{Nathaniel S. Pope}
\author[2,19,*]{Murillo F. Rodrigues}
\author[2,*]{Silas Tittes}
\author[3,*]{Linh N. Tran}

% additional authors
\author[4]{Ornob Alam}
\author[5]{Maria Izabel A. Cavassim}
\author[6]{Peter D. Fields}
\author[7]{Benjamin C. Haller}
\author[8,9]{Xin Huang}
\author[10]{Ben Jeffrey}
\author[2]{Kevin Korfmann}
\author[5]{Christopher C. Kyriazis}
\author[2]{Jiseon Min}
\author[11]{Inés Rebollo}
\author[2]{Clara T. Rehmann}
\author[2]{Scott T. Small}
\author[2]{Chris C. R. Smith}
\author[12]{Georgia Tsambos}
\author[10]{Yan Wong}
\author[13]{Yu Zhang}

% PI additional authors
\author[14]{Christian D. Huber}
\author[13]{Gregor Gorjanc}
\author[15]{Aaron P. Ragsdale}

% PIs that were very involved in the project
\author[16]{Ilan Gronau}
\author[3]{Ryan N. Gutenkunst}
\author[10]{Jerome Kelleher}
\author[5]{Kirk E. Lohmueller}
\author[17]{Daniel R. Schrider}
\author[2,18]{Peter L. Ralph}
\author[2]{Andrew D. Kern}

\affil[*]{\small{These authors contributed equally to the paper.}}
\affil[1]{\small{Section for Molecular Ecology and Evolution, Globe Institute, University of Copenhagen, Denmark}}
\affil[2]{\small{Institute of Ecology and Evolution, University of Oregon, Eugene OR 97402, USA}}
\affil[3]{\small{Department of Molecular and Cellular Biology, University of Arizona, Tucson, AZ 85721, USA}}
\affil[4]{\small{Center for Systems Biology, New York University, New York NY, 10012 USA}}
\affil[5]{\small{Department of Ecology and Evolutionary Biology, University of California, Los Angeles, Los Angeles CA, USA}}
\affil[6]{\small{The Jackson Laboratory, Bar Harbor, ME, 04609 USA}}
\affil[7]{\small{Department of Computational Biology, Cornell University, Ithaca NY, USA}}
\affil[8]{\small{Department of Evolutionary Anthropology, University of Vienna, Vienna, Austria}}
\affil[9]{\small{Human Evolution and Archaeological Sciences (HEAS), University of Vienna, Vienna, Austria}}
\affil[10]{\small{Big Data Institute, Li Ka Shing Centre for Health Information and Discovery, University of Oxford, Oxford OX3 7LF, UK}}
\affil[11]{\small{Instituto Nacional de Investigación Agropecuaria (INIA), Estación Experimental Las Brujas, Ruta 48 km 10, Canelones, Uruguay}}
\affil[12]{\small{Department of Genome Sciences, University of Washington, Seattle WA, USA}}
\affil[13]{\small{The Roslin Institute and Royal (Dick) School of Veterinary Studies, University of Edinburgh, Edinburgh EH25 9RG, UK}}
\affil[14]{\small{Department of Biology, Pennsylvania State University, University Park, PA 16802, USA}}
\affil[15]{\small{Department of Integrative Biology, University of Wisconsin-Madison, Madison WI, USA}}
\affil[16]{\small{Efi Arazi School of Computer Science, Reichman University, Herzliya 4610101, Israel}}
\affil[17]{\small{Department of Genetics, University of North Carolina at Chapel Hill, Chapel Hill NC 27599, USA}}
\affil[18]{\small{Department of Data Science, University of Oregon, Eugene OR 97402, USA}}
\affil[19]{\small{Division of Genetics, Oregon National Primate Center, Oregon Health and Science University, Beaverton OR 97006, USA}}

@@

    \subsection*{Inference of $N_e(t)$}
    Inference of $N_e(t)$ was conducted on simulated datasets after masking non-recombining portions
    of the genome (to avoid bias caused by extreme recombination rate variation)
    and exons (to avoid the direct effect of natural selection).
    Thus, inference from datasets simulated with selection was not influenced by direct selection
    acting on exons, but it was influenced by selection acting on linked sites (background selection).    
    %
    We assessed the accuracy of demographic inferences from three software packages:
    \msmc \citep{Schiffels2020}, \stairway \citep{liu2020stairway}, \gone \citep{santiago2020recent},
    and \smcpp \citep{terhorst2017robust}.
    For \msmc %\citep{Schiffels2020}, IG: no need to double-cite
    we used a random sample of $N=6$ simulated genomes with 20
    iterations of the EM algorithm.
    For \gone we set \texttt{max\_snps} to 500,000,
    the number of generations to 2,000, and the number of bins to 400, all representing default settings.
    For \stairway we did one run of $\theta$ estimation for each replicate dataset,
    and used a \texttt{dimFactor} of 5000, both of which represent default settings.
    For \smcpp we used the unfolded SFS (\texttt{--unfold}), 
    along with the true mutation rate under which the simulations were generated. 
    The complete \snakemake workflow for demographic inference can be found at
    \url{https://github.com/popsim-consortium/analysis2/blob/main/workflows/n_t.snake}.

>>>>

% author tranche explanation:
% co-first authors
\author[1,*]{Graham Gower}
\author[2,*]{Nathaniel S. Pope}
\author[2,3,*]{Murillo F. Rodrigues}
\author[2,*]{Silas Tittes}
\author[4,*]{Linh N. Tran}

% additional authors
\author[5]{Ornob Alam}
\author[6]{Maria Izabel A. Cavassim}
\author[7]{Peter D. Fields}
\author[8]{Benjamin C. Haller}
\author[9,10]{Xin Huang}
\author[11]{Ben Jeffrey}
\author[2]{Kevin Korfmann}
\author[6]{Christopher C. Kyriazis}
\author[2]{Jiseon Min}
\author[12]{Inés Rebollo}
\author[2]{Clara T. Rehmann}
\author[2]{Scott T. Small}
\author[2]{Chris C. R. Smith}
\author[13]{Georgia Tsambos}
\author[11]{Yan Wong}
\author[14]{Yu Zhang}

% PI additional authors
\author[15]{Christian D. Huber}
\author[14]{Gregor Gorjanc}
\author[16]{Aaron P. Ragsdale}

% PIs that were very involved in the project
\author[17]{Ilan Gronau}
\author[4]{Ryan N. Gutenkunst}
\author[11]{Jerome Kelleher}
\author[6]{Kirk E. Lohmueller}
\author[18]{Daniel R. Schrider}
\author[2,19]{Peter L. Ralph}
\author[2]{Andrew D. Kern}

\affil[*]{\small{These authors contributed equally to the paper.}}
\affil[1]{\small{Section for Molecular Ecology and Evolution, Globe Institute, University of Copenhagen, Denmark}}
\affil[2]{\small{Institute of Ecology and Evolution, University of Oregon, Eugene OR 97402, USA}}
\affil[3]{\small{Division of Genetics, Oregon National Primate Center, Oregon Health and Science University, Beaverton OR 97006, USA}}
\affil[4]{\small{Department of Molecular and Cellular Biology, University of Arizona, Tucson, AZ 85721, USA}}
\affil[5]{\small{Center for Systems Biology, New York University, New York NY, 10012 USA}}
\affil[6]{\small{Department of Ecology and Evolutionary Biology, University of California, Los Angeles, Los Angeles CA, USA}}
\affil[7]{\small{The Jackson Laboratory, Bar Harbor, ME, 04609 USA}}
\affil[8]{\small{Department of Computational Biology, Cornell University, Ithaca NY, USA}}
\affil[9]{\small{Department of Evolutionary Anthropology, University of Vienna, Vienna, Austria}}
\affil[10]{\small{Human Evolution and Archaeological Sciences (HEAS), University of Vienna, Vienna, Austria}}
\affil[11]{\small{Big Data Institute, Li Ka Shing Centre for Health Information and Discovery, University of Oxford, Oxford OX3 7LF, UK}}
\affil[12]{\small{Instituto Nacional de Investigación Agropecuaria (INIA), Estación Experimental Las Brujas, Ruta 48 km 10, Canelones, Uruguay}}
\affil[13]{\small{Department of Genome Sciences, University of Washington, Seattle WA, USA}}
\affil[14]{\small{The Roslin Institute and Royal (Dick) School of Veterinary Studies, University of Edinburgh, Edinburgh EH25 9RG, UK}}
\affil[15]{\small{Department of Biology, Pennsylvania State University, University Park, PA 16802, USA}}
\affil[16]{\small{Department of Integrative Biology, University of Wisconsin-Madison, Madison WI, USA}}
\affil[17]{\small{Efi Arazi School of Computer Science, Reichman University, Herzliya 4610101, Israel}}
\affil[18]{\small{Department of Genetics, University of North Carolina at Chapel Hill, Chapel Hill NC 27599, USA}}
\affil[19]{\small{Department of Data Science, University of Oregon, Eugene OR 97402, USA}}

@@

    \subsection*{Inference of $N_e(t)$}
    Inference of $N_e(t)$ was conducted on simulated datasets after masking non-recombining portions
    of the genome (to avoid bias caused by extreme recombination rate variation)
    and exons (to avoid the direct effect of natural selection).
    Thus, inference from datasets simulated with selection was not influenced by direct selection
    acting on exons, but it was influenced by selection acting on linked sites (background selection).    
    %
    We assessed the accuracy of demographic inferences from three software packages:
    \msmc \citep{Schiffels2020}, \stairway \citep{liu2020stairway}, \gone \citep{santiago2020recent},
    and \smcpp \citep{terhorst2017robust}.
    For \msmc 
    we used a random sample of $N=6$ simulated genomes with 20
    iterations of the EM algorithm.
    For \gone we set \texttt{max\_snps} to 500,000,
    the number of generations to 2,000, and the number of bins to 400, all representing default settings.
    For \stairway we did one run of $\theta$ estimation for each replicate dataset,
    and used a \texttt{dimFactor} of 5000, both of which represent default settings.
    For \smcpp we used the unfolded SFS (\texttt{--unfold}), 
    along with the true mutation rate under which the simulations were generated. 
    The complete \snakemake workflow for demographic inference can be found at
    \url{https://github.com/popsim-consortium/analysis2/blob/main/workflows/n_t.snake}.
